Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-11-11 eta.tex
Indicate master file in appendices for easier compilation

\documentclass[usenatbib,fleqn]{mn2e}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{epstopdf}
\epstopdfsetup{outdir=./figures/}
\graphicspath{{./figures/}}
\usepackage{url}
\usepackage{aas_macros}
\usepackage{astro}
% \usepackage{natbib}

% \renewcommand{\l}{\left}
% \renewcommand{\r}{\right}

\newcommand{\Mdot}{\dot{M}}
\newcommand{\eddr}{\dot{M}/\dot{M}_{\rm Edd}}
\newcommand{\Mdote}{\dot{M}_{\rm Edd}}
\newcommand{\Mdotb}{\dot{M}_{\rm bondi}}

\newcommand\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$<$}}}
\newcommand\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
    \raise1pt\hbox{$>$}}}
\newcommand{\rs}{r_s}
\newcommand{\rb}{r_b}
\newcommand{\rcirc}{r_{\rm circ}}
\newcommand{\rss}{r_{\rm ss}}
\newcommand{\lrs}{\l_{\rs}}
\newcommand{\lambdars}{\lambda_{\rs}}
\newcommand{\vw}{v_w}

\newcommand{\dxdy}[2]{\frac{d #1}{d #2} }
\newcommand{\ddr}[1]{\dxdy{#1}{r}}
\newcommand{\drhodt}{\dxdy{\rho}{t}}
\newcommand{\dpdr}{\dxdy{p}{r}}
\newcommand{\dvdr}{\dxdy{v}{r}}
\newcommand{\dsdr}{\dxdy{s}{r}}
\newcommand{\dphidr}{\dxdy{\Phi}{r}}

\newcommand{\ke}{\frac{v^2}{2}}
\newcommand{\kew}{\frac{v_w^2}{2}}

\newcommand{\gammaf}{\frac{\gamma}{\gamma-1}}
\newcommand{\gammafi}{\frac{\gamma-1}{\gamma}}
\newcommand{\cs}{\frac{p}{\rho}}
\newcommand{\Q}{q (\ke+\kew-\gammaf \cs)}

\newcommand{\kb}{k_{\rm b}}
\renewcommand{\mp}{m_{\rm p}}
\newcommand{\pc}{\rm pc}

\newcommand{\Menc}{M_{\rm enc}}
\newcommand{\rhostar}{\rho_*}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\Mseight}{M_{\star,8}}
\newcommand{\Mbh}[1][]{M_{\bullet#1}}
\newcommand{\Mbheight}{M_{\bullet,8}}
\newcommand{\MbheightExp}{\frac{\Mbh}{10^8 \Msun}}

\newcommand{\phirs}{\frac{G \Menc}{\rs}}
\newcommand{\soi}{\rm soi}
\newcommand{\rsoi}{r_{\soi}}
\newcommand{\ff}{\rm ff}
\newcommand{\tff}{t_{\ff}}
\newcommand{\rIa}{r_{\rm Ia}}
\newcommand{\EIa}{E_{\rm Ia}}
\newcommand{\RateIa}{R_{\rm Ia}}
\newcommand{\sigsoi}{\sigma_{\soi}}

\newcommand{\vwO}{v_{w,0}}
\newcommand{\kewO}{\frac{\vwO^2}{2}}
\newcommand{\x}{\frac{r_s}{\rsoi}}
\newcommand{\vwNorm}{\frac{\vwO}{\sigsoi}}
\newcommand{\vwOFH}{v_{w,0,500}}
\newcommand{\vwOFHexp}{\frac{\vwO}{500 \, {\rm km/s}}}

\newcommand{\pyear}{{\rm yr}^{-1}}
\newcommand{\tage}{t_{\star}}
\renewcommand{\th}{t_h}

\topmargin -1 cm
\defcitealias{WangMerritt:2004a}{WM04}

\author[Generozov, Stone, \& Metzger]{Aleksey Generozov$\thanks{E-mail: ag@astro.columbia.edu}$, Nicholas Stone, Brian~D.~Metzger\\
  Columbia Astropysics Labratory, Columbia University, 550 West 120th Street, New York, NY 10027}

\begin{document}
\title{Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes}
\maketitle

\begin{abstract}
  We calculate the steady-state hydrodynamic profiles of hot gas in
  the nuclei of a sample of early-type quiescent galaxies.  We assume
  that mass is supplied to the circumnuclear medium (CNM) entirely by
  stellar winds according to the radial stellar profiles for each
  galaxy, while the gas is heated by a combination of shocked stellar
  winds and supernovae.  Our analytic and numerical results are used
  to estimate the accretion rates onto the central supermassive black
  holes (SMBH) as a function of the SMBH mass $M_{\bullet}$ and the
  heating efficiency, the latter of which is related schematically to
  the age of the stellar population, the frequency of supernovae, and
  the efficacy of accretion feedback processes.  For a fixed heating
  efficiency, we find that the SMBH accretion Eddington factor
  increases with the SMBH mass, $\dot{M}/\dot{M}_{\rm edd} \propto
  M_{\bullet}^{\alpha}$ with $\alpha \approx 0.5, 1$ for cusp and core
  nuclei, respectively.
  % However, considering the potentially lower heating efficiency of low
  % mass SMBHs (due to a younger stellar population and the rareity of
  % supernovae within the stagnation radius) and their resulting higher
  % susceptibility to cooling instability, we find a flatter
  % relationship $\alpha \approx 0$. 
  Our results are used to check previous estimates of the SMBH
  accretion rate assuming Bondi accretion and extrapolating the X-ray
  measured density and temperature to the unresolved Bondi radius.  We
  also discuss the implications of our results for the diversity of
  CNM environments encountered by the jetted outflows from tidal
  disruption events, and for the use of an assumed
  $\dot{M}-M_{\bullet}$ relationship to constrain the occupation
  fraction of SMBHs in low mass galaxies.
\end{abstract}

\begin{keywords}
  black hole physics --  galaxies: active
\end{keywords}

\section{Introduction}
\label{sec:introduction}

%% Maybe active galaxies constitute 10% need reference here?
Supermassive black holes (SMBHs) lurk in the centers of most, if not
all nearby galaxies (see reviews by,
e.g. \citealt{KormendyRichstone:1995a};
\citealt{FerrareseFord:2005a}). However, only a few percent of these
manifest themselves as luminous Active Galactic Nuclei (AGN).  Nearly
quiescent SMBHs, such as those hosting low luminosity AGN, constitute
a silent majority (e.g.~\citealt{Ho:2009a}).

In order to understand why most SMBHs appear to be inactive, it is
necessary to understand their gaseous environments.  Hot gas near the
SMBH sphere of influence (the `circumnuclear medium' or CNM) controls
the central mass accretion rate $\dot{M}$.  Better knowledge of how
$\dot{M}$ varies with the SMBH mass $\Mbh$ and other properties of the
galactic nucleus is necessary to understand the co-evolution of SMBHs
and their host galaxies with cosmic time.  SMBH growth in the low
redshift universe, for instance, appears to be dominated by low mass
black holes (e.g.~\citealt{Heckman+04}).  Although this is often
attributed to cosmological ``down sizing" (e.g, \citealt{Gallo+08}),
the physical processes involved may be distinct from those operating
at higher SMBH masses and accretion rates in quasars.

The accretion rates of low mass SMBHs and their resulting nuclear
X-ray luminosities $L_X$ also have implications for observational
constraints on their occupation fraction in low mass galaxies and the
possible existence of intermediate mass black holes.
\citet{MillerGallo+:2014a} combine the relationship between $\Mbh$ and
$L_X$ from a sample of relatively high mass galaxies
(\citealt{Gallo+08}) with upper limits on $L_X$ in several lower mass
galaxies, to infer tentative evidence for an occupation fraction less
than unity.  However, this method relies on extrapolating a power-law
fit to the $L_X-\Mbh$ distribution to low $\Mbh$, below the
instrumental detection threshold.  The conclusions of
\citet{MillerGallo+:2014a} are thus sensitive to possible changes in
the $L_X$-$\Mbh$ relationship at small $\Mbh$.
% AG: In our model we obtain a steeper relationship, but we have difficulty matching the observed relationship above the detection threshold

The density profile of the CNM also influences the emission from
stellar tidal disruption events (TDEs), such as the high energy
transient {\it Swift} J1644+57 (\citealt{Levan+11};
\citealt{Bloom+11}, \citealt{Burrows+11}; \citealt{Zauderer+11}).
This event was powered by a transient relativistic jet, which produced
synchrotron radio emission as the jet material was decelerated by
shock interaction with the CNM of the previously quiescent SMBH
(\citealt{Giannios&Metzger11}; \citealt{Zauderer+11}).  Detailed
modeling of {\it Swift} J1644+57 showed that the CNM density was
significantly lower than that measured surrounding Sgr A$^{\star}$
(\citealt{Metzger+12}; \citealt{Berger+12}).  However, a TDE jet
injected into a denser CNM would be decelerated more rapidly,
producing substantially different radio emission than in {\it Swift}
J1644+57.  Variations in the properties of the CNM could thus in
principle help explain why most TDEs appear to be radio quiet
(e.g.~\citealt{Bower+13}; \citealt{VanVelzen+13}).

Gas that forms the CNM of a SMBH may originate from several sources
(\citealt{Ho:2009a}): (1) wind mass loss from (predominantly evolved)
stars; (2) star-star collisions; (3) the unbound debris from TDEs.
Stellar wind mass loss is probably the dominant source in most cases
insofar as collisions are relevant only in very dense stellar
environments for young stellar populations (\citealt{Rubin&Loeb11}),
while TDEs are estimated to provide a subdominant contribution to the
time-averaged accretion rate of quiescent SMBHs
(\citealt{MacLeod+13}).

\citet{Ho:2009a} calculates the accretion rates onto a sample of
early-type galaxies both by using X-ray observations to determine the
Bondi accretion rate, and by estimating the contributions from the
mass loss of evolved stars.  Both methods lead him to conclude that
the available reservoir of gas is more than sufficient to power
low-luminosity AGN, assuming the standard radiative efficiency $\eta
\sim 0.1$ for thin disk accretion.  Indeed, there is now strong
observational evidence that the low levels of AGN activity in
quiescent galaxies results from accretion proceeding in a radiatively
inefficient mode (\citealt{Yuan&Narayan14}), either due to the
gravitationally-released energy being advected across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to only a fraction of the gas
inflowing on large scales ultimately reaching the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

The accretion rates of SMBHs can in principle be estimated directly by
assuming Bondi accretion and using {\it Chandra} X-ray observations to
estimate the gas temperature and density near the Bondi radius.  This
technique has led to the intriguing result that the Bondi accretion
power correlates with the AGN jet power (e.g.,
\citealt{AllenDunn+:2006a}; \citealt{Russell+13};
\citealt{FujitaKawakatu+:2014a}), the latter estimated from the energy
required to inflate the observed cavities within the X-ray emitting
gas.  It is, however, almost never possible to resolve the Bondi
radius, thus requiring all such studies to extrapolate the observed
gas temperature and density profiles to much smaller radial scales.
%In fact, the temperature profile is assumed to be flat inside of the
%Bondi radius.  In reality, the cusp in the velocity dispersion near
%the SMBH, should cause a cusp in the gas temperature profile (assuming
%the kinetic energy of stellar winds is efficiently thermalized in
%shocks).
%% AG-give sense of scale.

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected to the nuclear environment via
stellar winds, while energy input results from a number of sources,
such as wind shock heating and supernovae (e.g.,
\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, however, which focus on modeling the nuclei of
individual galaxies, here we systematically analyze the properties of
the CNM across a large sample of quiescent galaxies with a range of
SMBH masses and other properties (\citealt{WangMerritt:2004a};
hereafter WM04).  For each galaxy, stellar winds deposit mass and
energy across a range of radii calculated directly from the observed
{\it stellar} density profile.

The one-dimensional steady state flow of gas is usually characterized
by an inflow-outflow structure, with a critical radius known as the
``stagnation radius" $\rs$ where the radial velocity goes to zero.
Mass loss from stars interior to the stagnation radius is accreted,
while that outside $\rs$ is unbound in an outflow from the nucleus.
The stagnation radius, rather than the Bondi radius, is thus
ultimately responsible for setting the SMBH accretion rate, although
$\rs$ can in many cases lie within a factor of few of the
nominal Bondi radius.
% Furthermore, in some cases we find that no stagnation
% radius exists at all; this formally implies that the entire of ISM of
% the galaxy is feeding the SMBH, potentially resulting in much higher
% accretion rate than would be predicted by simple Bondi accretion. 
The main goal of our analysis is to provide a framework for
interpreting the properties of quiescent galactic nuclei given the
comparatively well understood physics of stellar winds and supernova
heating.

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:sample}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe our
results.  In $\S\ref{sec:applications}$ we describe applicaiton of our
results, including a comparison of our gas profiles to those inferred
from {\it Chandra} observations and to the measured $L_X-\Mbh$
relationship.  In $\S\ref{sec:discussion}$ we discuss our results.  In
$\S\ref{sec:summary}$ we summarize our results.
\section{Model}
\label{sec:model}

\subsection{Galaxy Sample}

\label{sec:sample}
%AG:It is unclear whether if/how the Nuker profile will show up in
%this parametrized study.
Our sample includes 61 early-type galaxies from
\citetalias{WangMerritt:2004a}, as represent a subset of a larger
sample of \citet{FaberTremaine+:1997a} with Hubble WFPC2 imaging.  We
focus on elliptical galaxies because we are interested in isolating
SMBHs fed by the steady mass loss from the relatively old stellar
population, instead of fresh gas flowing into the sphere of influence
from larger scales.

The surface brightness of each galaxy in the sample is parameterized by a Nuker Law (WM04; their Table 1),
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b}.
\end{equation}
The Nuker profile is a broken power law which transitions from an inner power law slope, $\Gamma$, to an outer power law slope, $\beta$, at a break radius, $\rb$.

Given $I(\xi)$ we follow the procedure outlined by
\citetalias{WangMerritt:2004a} to compute the spherically-symmetric
stellar density profile $\rhostar(r)$ and stellar mass profile
$M_{\star}(r) = 4\pi \int \rhostar r^{2}dr$, both of which are used in
our model.  For reference, note that stellar density $\rhostar\sim
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\sim r^{-1-\beta}$ for $r
\gg \rb$.  Since $\Gamma\approx 1$ for ``cusp" galaxies, while
$\Gamma\approx 0$ for ``core" galaxies, at small radii $\ll \rb$ we
thus have $\rhostar \propto r^{-\alpha}$ and $M_{\star} \propto
r^{\beta}$ with $\beta = 3-\alpha \approx 1-2$.  For each galaxy the
SMBH mass $\Mbh$ is determined from the $\Mbh-\sigma$ relationship.
%%AG-ugh I just took the Mbh from Wang & Merritt, which I guess is
%%calculated from an outdated M-sigma. It would be bizarre to use 2
%%different relations.

% AG:Is there a systematic dependence of beta on galaxy type$.

%% AG--some discussion of alternate parameterizations.  My intuition
%% is that for our purposes the parameterization should not
%% matter. But some discussion of alternate possibilities could
%% prevent Alistar Graham from yelling at us.

\subsection{Hydrodynamic Equations}
\label{sec:hydro}

Following \citet{Quataert:2004a} (see also \citealt{HolzerAxford:1970a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}), we calculate the gas density profile $\rho$ of the CNM for each galaxy within our sample by solving the equations of one-dimensional (spherically symmetric) time-dependent hydrodynamics,
\begin{align}
  &\drhodt+\frac{1}{r^2}\frac{\partial}{\partial r}\left(\rho r^2 v\right)=q \label{eq:drhodt}\\
  &\rho \left(\frac{\partial v}{\partial t} + v\frac{\partial
      v}{\partial r}\right) =-\frac{\partial p}{\partial r}- \rho\frac{GM_{\bf enc}}{r^{2}} -q v \label{eq:dvdt}\\
  &\rho T\left(\frac{\partial s}{\partial t} + v\frac{\partial
      s}{\partial r}\right)=q\left[\ke+\kew-\gammaf \cs \right] +\nabla\cdot(\kappa \nabla T)
, 
\label{eq:dsdt}
\end{align}
where $v$, $T$, $p$ and $s$ are the radial velocity, temperature,
pressure and specific entropy of the gas, respectively, and $M_{\rm enc} = M_{\star}(r) + \Mbh$ is the
enclosed mass.  We assume an ideal gas equation of state with $p =
\rho kT/\mu m_p$ with $\mu = 0.62$  and $\gamma = 5/3$. {\bf BDM: take $\mu = 0.62$ here for fully ionized gas}

The source term in the continuity equation (equation~\ref{eq:drhodt}),
%Should equations be referenced as equation of equation
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 10^{10}$ years (see, e.g.,
\citealt{Ciotti+91}).  In Appendix \ref{app:eta} we estimate that
$\eta \approx 0.11(t_{\star}/t_{\rm h})^{-1.26}$ for a Salpeter IMF,
assuming that all stars are formed in a burst of star formation at a
common age of $t_{\star}$ (equation~[\ref{eq:eta}]).  We adopt
$\eta=0.1$ as fiducial {\bf BDM: take $\eta = 0.1$ instead of 1}, but note that changing $\eta$ is equivalent to
rescaling $\rho$ by the same factor.\footnote{This is {\it not} true
  if cooling is included, since the cooling rate is proportional to
  density {\it squared}.}

Source terms $\propto q$ also appear in the momentum and entropy
equations (equation~[\ref{eq:dvdt}] and [\ref{eq:dsdt}]) due to the fact
that the isotropic mass injection in the rest frame of the SMBH
represents a source of momentum and energy relative to the mean flow (see also below).  
%No net momentum is in the winds, but mass is being added which adds a
%term in the velocity equation...

The second source term in equation (\ref{eq:dsdt}) accounts for heating transportation by electron conductivity, where the conductivity $\kappa = \kappa_0 T^{5/2}$ is assumed to equal the classical ``Spitzer" value (\citealt{Spitzer62}); although even a weak magnetic field could in principle strongly suppress the conductivity perpendicular to the field, the radially-decreasing temperature profiles of interest are susceptible to the magneto-thermal instability (\citealt{Balbus01}), the non-linear evolution of which is believed to result in a radially-directed field structure (\citealt{Parrish&Stone07}).

Following \citet{ShcherbakovWong+:2014a}, the term in the entropy
equation $\propto \vw^2 = \sigma(r)^2+v_{w,0}^2$ accounts for external
heating sources.  Here $\sigma(r) = \sqrt{GM_{\rm enc}/r}$ is the
velocity dispersion of the stars, thus accounting for the fact that at
a minimum the gas experiences shock heating between the stellar winds
of velocities moving with respect to each other a characteristic
velocity $\sim \sigma$.  This heating is present if the wind velocities themselves
are low, $\ll \sigma$, as characterize the majority of AGB mass loss.
The term $v_{w,0}^{2}$ parameterizes additional sources of heating,
such as main sequence winds, millisecond pulsars, supernovae, etc.  As
discussed in $\S\ref{sec:heating}$, the value of $v_{w,0}$ is
uncertain and will in general depend on the SMBH mass and the age of
the stellar population.

Equations (\ref{eq:drhodt})-(\ref{eq:dsdt}) are solved using a sixth
order finite difference scheme with a third order Runge-Kutta scheme
for time integration and artificial viscosity terms in the velocity
and entropy equations for numerical stability (\citealt{Brandenburg:2003a}).  We solve the equations for
different choices of $v_{\rm w,0}$ spanning the physically expected
range.  Although we are most interested in the steady-state CNM
structure, the time-dependent equations are solved to avoid numerical
issues that arise near the critical sonic points, of which there are
up to two because of the supersonic inflow on small radial scales and
a possible supersonic outflow on large scales.

We do not include radiative cooling in the entropy equation because
across much of the parameter space radiative cooling is negligible
compared to other sources of cooling (e.g., advection).  Although
radiative cooling cannot be neglected in some cases (especially if
$\vwO$ is low or $\eta$ is high), our 1D scheme cannot account for the
true multi-phase behavior of local thermal instabilities anyways
(e.g.~\citealt{McCourt+12}).  Our steady state solutions neglecting
radiation can nevertheless be used to ascertain over what parameter
space cooling instabilities are likely to develop
($\S\ref{sec:cooling}$).
%Dependence of the heating to cooling ratio has a far steeper
%dependence on vwO than on eta...

The accuracy of our numerical solutions is assessed by checking the
level of conservation of mass across the grid, as well as the integral
constraint on the energy.  Mass and energy are conserved to $\lesssim
40$ per cent for all solutions presented in this
paper. 
{\bf BDM: more details...}

\section{Results}

We begin by describing a number of useful analytic results before describing our numerical solutions.

\label{sec:results}
\subsection{Analytics: Flow properties at $\rs$}

This section summarizes analytic estimates of several physical quantities of interest, such as the stagnation radius $\rs$ and the mass accretion rate, the detailed derivation of which are given in Appendix \ref{app:rs}.  All of our results are derived under the assumption that thermal conductivity can be neglected.

When possible our results are expressed in terms of the wind heating
parameter $v_{\rm w,0}$; the power-law slope of the density profile
$n$ (also evaluated at $r = \rs$); and the dimensionless parameter \be
\chi \equiv v_{\rm w}^{2}/v_{\rm w,0}^{2} \approx \left(1 + \sigma^{2}/v_{\rm w,0}^{2}\right) \approx
\left(1+0.14 \Mbheight^{0.43}v_{500}^{-1}\right), \ee where $v_{500}
\equiv v_{w,0}/500$ km s$^{-1}$, $M_{\bullet,8} \equiv
M_{\bullet}/10^{8}M_{\odot}$, and the second equality makes use of the
$M_{\bullet}-\sigma$ correlation (\citealt{Gultekin+09}) \be
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\ee
Although the use of equation (\ref{eq:Msigma}) is questionable for low mass BHs (e.g., \citealt{Greene&Ho07}), our results are not particularly sensitive to this assumption.

First, by demanding continuity of the entropy derivative at the stagnation radius, the temperature at the stagnation radius is given by.
\begin{align}
T \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p v_{w}^{2}}{2} \approx 3.7
\times 10^6\,\chi v_{500}^2 \,\,{\rm K} 
\label{eq:Tanalytic}
\end{align}
%{\bf BDM: i picked $\mu = 0.62$ and stick to it.}

Demanding hydrostatic equilibrium at the stagnation point provides an expression for the stagnation radius
\begin{align}
  \x=\frac{1}{\zeta^2 n}\left[
    \frac{\Mstar}{\Mbh}\left(A-\frac{n}{2}\right)+(B-\frac{n}{2})\right],
  \label{eq:stag_analytic}
\end{align}
where $\zeta \equiv v_{w,0}/\sigma$, $n$ is the power-law slope of the
density profile at $r = r_{\rm s}$,

\begin{align}
r_{\rm soi} = \frac{GM_{\bullet}}{\sigma^{2}} \approx 14M_{\bullet,8}^{0.6}\,\,{\rm pc}
\end{align}

is the radius of the SMBH sphere of
influence (soi), and A and B are constants that depend on the Nuker
parameter $\Gamma$ and the adiabatic index of the gas.

If the SMBH dominates the potential at the stagnation radius
($M_{\bullet} \gg M_{\star}$; $v_{w,0} \gg 200 M_{\bullet,8}^{0.2}$ km
s$^{-1}$) then the relationship between $\rs$ and $v_{w}$ is
simplified greatly:
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n \vwO^2} \simeq 6 \, \pc \,\, \Mbheight v_{500}^{-2},
  \label{eq:stag_simple}
\end{align}
where in the numerical estimate we have assume $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on heuristic grounds (their equation~[6]).  On the other hand, if wind heating is much weaker ($v_{\rm w,0} \ll \sigma$), the stagnation radius moves to very large radii or may not exist at all.

The importance of conductive heating can be estimated by comparing its
magnitude to that of wind heating, viz.~
\begin{eqnarray} \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim& \frac{2t_{\rm h}\kappa_0
T^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}(r_{\rm s}) v_{\rm w}^{2}}
\nonumber \\ &\sim&
0.5\left(\frac{\eta}{0.1}\right)^{-1}\left(\frac{r_{\rm s}}{r_{\rm
soi}}\right)^{-\beta} v_{500}^{3}\chi^{3},
\end{eqnarray}

where we have approximated $\nabla^{2} \sim 1/r_{\rm st}^{2}$ and in the second equality have made use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}) and have approximated $4\pi r_{\rm s}^{3}\rho_{\star}(r_{\rm s}) \approx M_{\bullet}(r_{\rm s}/r_{\rm soi})^{\beta}$ with $\beta \approx 1-2$.  We thus see that conductive heating is important near the stagnation radius for high values of $v_{\rm w} \gtrsim 500$ km s$^{-1}$ and $\eta \gtrsim 0.1$, but that it is safely neglected to first order for lower heating rates.

The SMBH accretion rate is simply given by the integrated mass loss
rate interior to the stagnation radius (equation~[\ref{eq:q}]), 
%AG: Should we have separate cores and cusps? Or just write a general

\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc(\rs)}{\th} \nonumber \\
&\approx&
  \begin{cases}
    1.5 \times 10^{-3} M_{\bullet,8}^{1.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta \Msun \, \pyear& \text{core} \\
    2.6 \times 10^{-3} \Mbheight^{1.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}

The Eddington ratio $\eddr$ is given by

\begin{eqnarray}
\eddr&\approx&
  \begin{cases}
    6.6 \times 10^{-4} M_{\bullet,8}^{0.76}
    \chi^{-3.8} v_{500}^{-3.8}  \eta \Msun \, \pyear& \text{core} \\
    1.2 \times 10^{-3} \Mbheight^{0.48} 
    \chi^{-2.4} v_{500}^{-2.4}  \eta \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}

The density of gas at the stagnation radius $\rho(r_{\rm s})$ can be estimated by using equations (\ref{eq:stag_simple}) and (\ref{eq:mdot_analytic}) in conjunction with an alternative estimate of the mass accretion rate as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}(\rs) = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$, viz.
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}(\rs)},
  \label{eq:mdot_gas}
\end{align}
from which one derives 
\begin{align}
  \rho(\rs)\simeq
  \begin{cases}
    3 \times 10^{-24} \Mbheight^{-0.14} v_{500}^{-1}  \eta \,
    \, {\rm g \, cm^{-3}}& \text{core}\\
    6 \times 10^{-24}  \Mbheight^{-0.56} v_{500}    \eta \,\, {\rm g \,cm^{-3}} & \text{cusp}
  \end{cases}
  \label{eq:rhors}
\end{align}

\subsection{Comparison with Bondi Accretion}
The spherical accretion rate onto a point source from a surrounding medium of fixed sound speed $c_{\rm s} = (kT/\mu m_p)^{1/2}$ and density is given by (\citealt{Bondi52}) 
\begin{align}
  \dot{M}_{\rm B} =4\pi \lambda r_b^2 \rho(r_b) v_{\rm ff}(r_b),
\label{eq:bondi}
\end{align}
where $r_b \equiv GM/c_{s}^{2}$ is the ``Bondi radius", $v_{\rm ff} = (GM_{\bullet}/r)^{1/2}$ and $\lambda$ is a parameter of order unity.

Rewriting equation (\ref{eq:mdot_gas}) as
\begin{align}
  &\dot{M}\simeq\frac{(4 \pi/3) \rs^3 \rho(\rs)}{t_{\ff}} =\frac{4 \pi}{3} \rs^2 \rho(\rs) v_{\ff}(\rs)
\end{align}
we note the qualitative similarity between this expression and the Bondi formula (equation~[\ref{eq:bondi}]) with the exception that $\rb$ is replaced by $\rs$.  Using equation (\ref{eq:stag_simple})
\begin{align}
  \rs\approx\frac{7}{2}\frac{G \Mbh}{\vwO^2} \approx \frac{14}{3}\frac{GM_{\bullet}}{c_{s}^{2}},
  \label{eq:rs_simple}
\end{align}
where in the second line we have used the fact that $v_{w}^{2}/2 =
c_s^2/(\gamma-1)$ at $\rs$.  Thus, provided that a stagnation radius
exists which is interior to the sphere of influence radius, the
accretion rate predicted by our model is quite close to $\dot{M}_{\rm
  B}$.
 %({\bf BDM: Check - as this depends on
 % whether eq 10 is correct})

%% AG-Apparently even in the low heating case, it is still possible to
%% an outflow (from the analytic results above the gas density must be
%% quite steep in this case). However, at low heating we will probably
%% become subject to a cooling instablity.  AG-I believe that in the
%% case of core galaxy, the negative root of the quadratic should be
%% taken...

\subsection{Numerical Solutions}

Numerical solutions are calculated for each galaxy for three different values of the stellar wind heating parameter $v_{w,0}$ = 200, 500, 1000 km s$^{-1}$.

Figure~\ref{fig:profiles} shows three representative profiles of the density $\rho(r)$, temperature $T(r)$, and radial velocity $|v(r)|$.  As expected, the gas density increases towards the SMBH.  For some profiles (e.g. NGC1172 and NGC4478) the density $\rho$ follows a power law from the stagnation radius (squares in Fig.~\ref{fig:profiles}) inwards $\rho \propto r^{-n}$ with $n \approx 1$. Note that this is shallower than the $-3/2$ power law for Bondi accretion. However, the power law behavior does not extend through all radii, and the gas density profile may have breaks mirroring those in the stellar density profile. For example, NGC4365 has a prominent break in the gas density profile, near the stagnation radius. This coincides with the location
of the break in the stellar light profile ($\rb$). The temperature profile is relatively flat at large radii, but increases as $\propto
1/r^{k}$ interior to the sphere of influence, where $k$ is smaller than 1--the power expected for virialized gas.  The inwardly
directed velocity increases towards the hole approaching the expected free-fall rate $v \propto v_{\rm ff}\propto r^{-1/2}$ interior to the sonic point, and then increases outwards towards large radii.
% {\bf BDM: I made this up - in general we need
%   more explanation of what the plots are actually showing, with some
%   physical explanation.}
% AG -- things are not quite approaching Bondi on the inner grid. I
% don't have a good physical explanation yet...
% AG -- I am somewhat concerned about boundary effects in
% NGC3115. Perhaps the break in the gas density profile is due to
% boundary effects--I know that in this case the break is not
% associated with a break in the stellar density profile

Figure~\ref{fig:stag} (top panel) shows numerical results for the stagnation radius $r_{\rm s}$ where the radial velocity goes to zero (expressed in ratio to the radius of the SMBH sphere of influence $r_{\rm soi}$), with different colors representing different values of $v_{w,0}$.  Cusp and core galaxies are marked with square and triangles, respectively.  
%The solid line is a power law fit through the cusp galaxies, while the dotted line is a power law fit through the core galaxies.  {\bf BDM: I see no lines!}

>>>>

We now describe the rate of start formation $q$ varies with the age of the stellar population.

If we assume that a stellar population forms impulsively in the distant pass with IMF $\mu(m_*)$(with minimum mass $m_0$ and maximum mass $m_1$), then the surviving mass fraction at any future time $t$ is given by 
\begin{equation}
f(t) =\frac{ \int_{m_0}^{m_{\rm max}(t)} m_* \mu(m_*) dm_* }{ \int_{m_0}^{m_1} m_* \mu(m_*) dm_* },
\end{equation}
where 
\begin{equation}
m_{\rm max}(t) \approx 2.5M_\odot~ \left( \frac{t}{10^9~{\rm yr}} \right)^{-0.4}.
\end{equation}
For a Salpeter IMF $\mu(m_*) \propto m_*^{-2.35}$ with $m_0=0.1M_\odot$ and $m_1=100M_\odot$,
\begin{equation}
f_{\rm Sal}(t) = 1.098 - 0.490 \left(\frac{t}{10^{10}~{\rm yr}} \right)^{0.14}
\end{equation}
%{\bf NCS: we should probably use a Kroupa/Chabrier IMF, but this gets the ball rolling.}
If we approximate post-main sequence evolution as instantaneous and define $\lambda(m_*)$ as the fractional mass lost during all stages of stellar evolution, then the mass loss rate density
\begin{equation}
q(t) = \frac{\rho_*}{\bar{m}_*} \lambda(m_{\rm max}(t)) m_{\rm max}(t) \frac{df}{dt},
\end{equation}
where the mean stellar mass $\bar{m}_* = \int_{m_0}^{m_{\rm max}(t)} m_*\mu(m_*)dm_* \approx 0.3 M_\odot$.  Further approximating $\lambda(m_*)=0.5$, and using the Salpeter IMF once more, gives
\begin{equation}
q(t) = \frac{\rho_*}{\th} \times 0.11 \left(\frac{t}{10^{10}~{\rm yr}} \right)^{-1.26}.
\end{equation}
This is a specific, time-dependent definition of $\eta$:

\begin{equation}
\eta(t) (=0.11(t/t_{\rm h})^{-1.26})
\label{eq:eta}
\end{equation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "ms"
%%% End:
